{\bf {\Large Title....}}

### 2016-06-08 ident_manu.Rnw
some general ideas about questions

\begin{abstract}
\noindent 
\end{abstract}
\acresetall

@@

Overall objective: Bio-geo-chemical models of water quality in coastal environments are useful tools for management and policy-making. Significant time and resources are spent developing these models in system-specific contexts. The transferability or relative ease of downscaling these models to different systems could facilitate their use as tools for environmental management. This study will evaluate a simple bio-geo-chemical model that is the core unit of a larger spatio-temporal model of hypoxia on the Louisiana Shelf of Gulf of Mexico. The analysis will focus on use of the model to describe dissolved oxygen dynamics over time having implications for understanding factors that contribute to hypoxic conditions that are detrimental to aquatic resources.

Questions:

* Can the FishTank model that is currently parameterized for GOM conditions be downscaled and parameterized to smaller systems (using Weeks Bay as a test example)?
* What parameters have the greatest influence on dissolved oxygen as identified by local and global sensitivity analyses?
* What is the 'parameter space' (from sensitivity analysis) and 'identifiability' of the model parameters?
* What is the uncertainty in identifiable parameters (MCMC eval)?
* Does the analysis identify information that is consistent with current understanding of ecological factors that influenc system condition?  That is, does the model give the right results for the right reasons?
* Additional questions may include the following:
    * What initial conditions have the greatest effect on dissolved oxygen?
    * How does the model structure influence predictions of dissolved oxygen?  Does changing complexity influence results, e.g., one to many phytoplankton groups?
    * How does the assimilation of additional parameters (e.g., other state variables) during calibration influence the conclusions?

Methods:

1. Develop a programmatic infrastructure for FishTank that facilitates sensitivity analysis, parameterization, and additional evaluation.
2. Conduct local and global sensitivity analyses of dissolved oxygen to identify parameters that influence predictions using existing values for the Gulf of Mexico
3. Parameterize FishTank to Weeks Bay conditions using NERRS dissolved oxygen time series and values for initial conditions that approximate observed.  This will include a characterization of parameter identifiability.  Appropriate calibration and validation datasets must also be identified.
4. Use Markov chain Monte Carlo analyses to estimate parameter uncertainties.
5. Evaluate model results in the context of ecological properties that influence hypoxia, with implications for additional data needs or further development of bio-geo-chemical models.

@@

@@

@@

@@

>>>>

\begin{abstract}
\noindent Bio-geo-chemical models are useful tools in environmental sciences that can guide management and policy-making. Significant time and resources are spent developing these models in system-specific contexts. The optimization of model parameters to maximize precision, including transferability of these models to different systems, are fundamental concerns in the development and application of these tools. This study describes quantitative limitations of model complexity by contrasting numeric and ecological certainty.  We evaluate a simple bio-geo-chemical model that is the 1-dimensional unit of a larger spatio-temporal model of hypoxia on the Louisiana Shelf of Gulf of Mexico as an example. Results from analysis of the 1D model are used to infer larger trends in dissolved oxygen dynamics over time, having implications for understanding factors that contribute to environmental conditions that are detrimental to aquatic resources.  In particular, we focus on issues of parameter identifiability using global sensitivity analyses to provide quantitative descriptions of numerical constraints on model precision.  We argue that quantitative and ecological certainty in model calibration are often at odds and the practitioner must explicitly choose model components to optimize given tradeoffs between the two. We conclude that numerically optimal parameter sets are often small subsets of the complete parameter set for a model because of redundancies in the unique effects of paramater perturbations on model output.  As a result, use of a model for inference into ecological mechanisms of observed or predicted changes in response to management actions can be potentially misguided.  Although these concerns have been expressed in the literature, they are rarely explicitly addressed or included in model evaluations.  In addition to immediate implications for regional models, we provide a framework for describing the effects of parameter uncertainty and identifiability that can be applied to similar models to better inform environmental management.
\end{abstract}
\acresetall

@@

\begin{enumerate}
\item Simulation/biogeochemical/process-based models overview, contrast with statistical models
\item What models seek to provide - generality, precision, realism \cite{Levins66}, there is a tradeoff so models are 1) developed in partial independence and dependence on the world and theory, 2) function autonomously from both, or 3) represent both at the same time, from \cite{Morrison99}, cited in \cite{Ganju16}.  This is similar to the bias-variance tradeoff for statistical models, e.g., overparameterization of a model makes it very biased as it fits the data (the world) exactly
\item How is model performance/uncertainty evaluated regarding what they should provide - structural, observational, parameter \cite{Beck87}?
\item Parameter uncertainy as low-hanging fruit - can do post-hoc and from inner to outer level of complexity
\item Challenges related to uncertainty - similar to degrees of freedom, identifiability definition from \cite{Brun01} and need to evaluate identifiability \cite{Fasham06}
\item objectives of study, definition of `optimum'
\begin{enumerate}
\item What parameters have the greatest influence on dissolved oxygen as identified by local and global sensitivity analyses?
\item What is the 'parameter space' (from sensitivity analysis) and 'identifiability' of the model parameters?
\item Can we get a better estimate of identifiability other than 10-15 according to \cite{Brun01,Soetaert10}?
\item Can we reconcile differences in numeric uncertainty and ecological uncertainty, e.g., given a finite parameter space, what is an appropriate way to choose parameters given question of interest?  
\end{enumerate}
\end{enumerate}

@@

\subsection{Model description}

GGEM, FishTank, parameters and parameter categories, how they compare to other models

\subsection{Local sensitivity analysis}

Methods from \cite{Soetaert10}

@@

\subsection{Global sensitivity and identifiability}

@@

\subsection{Identifying subsets and testing calibration}

@@

Questions specific to GOM - what initial conditions are important? How many phytoplankton groups do we need (e.g., related to structural uncertainty)?

How does the assimilation of additional parameters (e.g., other state variables) during calibration influence the conclusions?

How does uncertainty translate to what a model should provide (generality v precision)?  The first step - find out what can be optimized but then do not overfit....

Need to talk about \cite{Flynn05}

\clearpage
\begin{singlespace}
\bibliographystyle{apalike_mine}
\bibliography{refs}
\end{singlespace}
\clearpage

### 2016-06-08 ident_manu.tex
some general ideas about questions

\begin{singlespace}
\title{{\bf {\Large Title....}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, John C. Lehrter$^1$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Emails: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}, \href{mailto:lehrter.john@epa.gov}{lehrter.john@epa.gov}}}
  \vspace{1in} 
  \\ Version Date:   Wed Jun 8 08:37:59 2016 -0500
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

\begin{abstract}
\noindent 
\end{abstract}
\acresetall

>>>>

\begin{singlespace}
\title{{\bf {\Large Title....}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, John C. Lehrter$^1$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Emails: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}, \href{mailto:lehrter.john@epa.gov}{lehrter.john@epa.gov}}}
  \vspace{1in} 
  \\ Version Date:   Wed Jun 8 09:02:46 2016 -0500
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

\begin{abstract}
\noindent Bio-geo-chemical models are useful tools in environmental sciences that can guide management and policy-making. Significant time and resources are spent developing these models in system-specific contexts. The optimization of model parameters to maximize precision, including transferability of these models to different systems, are fundamental concerns in the development and application of these tools. This study describes quantitative limitations of model complexity by contrasting numeric and ecological certainty.  We evaluate a simple bio-geo-chemical model that is the 1-dimensional unit of a larger spatio-temporal model of hypoxia on the Louisiana Shelf of Gulf of Mexico as an example. Results from analysis of the 1D model are used to infer larger trends in dissolved oxygen dynamics over time, having implications for understanding factors that contribute to environmental conditions that are detrimental to aquatic resources.  In particular, we focus on issues of parameter identifiability using global sensitivity analyses to provide quantitative descriptions of numerical constraints on model precision.  We argue that quantitative and ecological certainty in model calibration are often at odds and the practitioner must explicitly choose model components to optimize given tradeoffs between the two. We conclude that numerically optimal parameter sets are often small subsets of the complete parameter set for a model because of redundancies in the unique effects of paramater perturbations on model output.  As a result, use of a model for inference into ecological mechanisms of observed or predicted changes in response to management actions can be potentially misguided.  Although these concerns have been expressed in the literature, they are rarely explicitly addressed or included in model evaluations.  In addition to immediate implications for regional models, we provide a framework for describing the effects of parameter uncertainty and identifiability that can be applied to similar models to better inform environmental management.
\end{abstract}
\acresetall
